How is experience transformed into memory?

### 2018-09-01 paper/main.tex
edits to event seg figure, backing up work

\section*{Results}
To characterize the gists of the \textit{Sherlock} episode participants watched and their subsequent recountings of the episode, we used a topic model~\cite{BleiEtal03} to discover the latent thematic content.  \cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) scenes spanning the roughly 45 minute video.  This information included: a brief narrative description of what was happening; whether the scene took place indoors vs. outdoors; names of any characters on the screen; names of any characters who were in focus in the camera show; names of characters who were speaking; the location where the scene took place; the camera angle (close up, medium, long, etc.); whether or not background music was present; and other similar details (for a full list of annoted features see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the ``vocabulary'' for the topic model.  We then concatenated the sets of words across all features contained in overlapping 50-scene sliding windows, and treated each 50-scene sequence as a single ``document'' for the purposes of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 28 unique topics were sufficient to describe the time-varying content of the movie (see \textit{Methods}; Fig.~\ref{fig:schematic}).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06}, in that we sought to characterize how the thematic content of the data evolved over time.

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/analysis_schematic_bw}
\caption{\small \textbf{Schematic of the analysis approach.} For each scene in the video, text descriptions were generated by hand. Three exemplary time points are displayed here.  Below the video descriptions are text samples from an example participant's verbal recall transcript.  We fit a topic model to the moment-by-moment video text descriptions and transformed participants' verbal recall transcripts using this same model. The bar charts display the resulting topic model weights for the video (in blue) and recall (in green) for three example topic dimensions.}
\label{fig:schematic}
\end{figure}

The topics we found appeared heavily character-focused (e.g., the top-weighted word in each topic was nearly always a character) and could be roughly divided into themes related to primarily Sherlock-focused; primarily (John) Watson-focused (Watson is Sherlock's close confidant and assistant); and themes that involved Sherlock and Watson interacting (Fig.~\topics).  Several of the topics also appeared highly similar, which we hypothesized might allow us to distinguish between subtle narrative differences (e.g., if the distinctions between those overlapping topics were meaningful), but could also impair characterization of the narrative (e.g., if the distinctions between topics were not meaningful).

@@

% characterize gist of movie and recalls
% 1. movie: fit topic models (add note on dynamic topic models).  fig 1: what do the topics reflect (top words from each topic), example with a few scenes.
% 2. notice block diagonal structure, rather than smooth transitions; motivates HMM approach.  Note about selecting K
% 3. question: do recalls exhibit same structure?  apply topic model to recalls.  note that by projecting everything into a common topic space, the thematic content of each moment of movie/recall can be compared, even though the mix of words used to describe any given event or moment may not be consistent across people (or across annotations/recall)
% 4. recalls exhibit blocky structure, but at different scales.  motivates fitting K separately by person
% 5. match up events (correlations)
% 6. now we can use classic memory analyses to examine structure of recall.  motivate and reference PFR, lag-CRP, and SPC.  but this does not tell us about structure in the *content* of what people remembered, or whether people's recalls are capturing the hierarchical temporal structure of the original movie.
% 7. topic trajectories as a new way of characterizing recall.  assess agreement using rayleigh test.  also examine shapes of each subject's trajectory to show match.  key insight: everyone captures the general shape, but at different "sampling resolutions"
% 8. we can also project back onto the original (word) features to understand more about what any particular part of the trajectory reflects (wordle fig)
% 9. the above analyses tell us about how the trajectory of thematic content in the episode is both preserved and transformed in participants' memories.  we next carried out analyses to understand the neural bases of these phenomena.  analysis 1: which structures are sensitive (during viewing) to the original episode's topic trajectory?  answer: default mode.  anlaysis 2: which structure are sensitive (during viewing) to how the participants will later recall the episode?  answer: vmPFC, default mode.  (potential role of vmPFC in reflecting participant-specific memory transformations and processing)

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recall.} A depiction of our analysis pipeline. For all plots, darker colors indicate greater values and the range of each plot is 0-1.  \textbf{A.} A timepoints (1976) by topics (100) matrix representing the video stimulus.  Each row represents the most likely mixture of topics for a given timepoint (i.e., topic weights). Each column represents a different topic. \textbf{B.} A sentences (294) by topics (100) matrix representing participant \#13's recall. \textbf{C.} A viewing-time (1976) by viewing-time (1976) correlation matrix representing the correlation of each moment of the video model with every other moment of the video model. The yellow boxes represent `events' recovered by a hidden Markov model. \textbf{D.} A recall-time (294 sentences) by recall-time (294) correlation matrix for participant \#13. \textbf{E.} An events (34) by topics (100) matrix where each row represents the average topic vector for each event in the video model.  \textbf{F.} An events (27) by topics (100) matrix where each row represents the average topic vector for each event in participant \#13's recall model. \textbf{G.} A recall events (27) by video events (34) correlation matrix for participant \#13. The cells with a yellow border identify the video event with the highest correlation to a given recall event. \textbf{H.} A group averaged recall events (34) by video events (34) correlation matrix.  The cells with yellow borders are the video event with the highest correlation to a given average recall event.}
\label{fig:model}
\end{figure}

>>>>

\section*{Results}
To characterize the gists of the \textit{Sherlock} episode participants watched and their subsequent recountings of the episode, we used a topic model~\cite{BleiEtal03} to discover the latent thematic content.  \cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) scenes spanning the roughly 45 minute video.  This information included: a brief narrative description of what was happening; whether the scene took place indoors vs. outdoors; names of any characters on the screen; names of any characters who were in focus in the camera show; names of characters who were speaking; the location where the scene took place; the camera angle (close up, medium, long, etc.); whether or not background music was present; and other similar details (for a full list of annoted features see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and,'' ``or,'' ``but,'' etc.) across all features and scenes as the ``vocabulary'' for the topic model.  We then concatenated the sets of words across all features contained in overlapping 50-scene sliding windows, and treated each 50-scene sequence as a single ``document'' for the purposes of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 28 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the movie (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06}, in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize the how the properties of collections of documents change over time, our approach allows us to examine the topic dynamics within a single episode.  Specifically, our approach yielded a single \textit{topic vector} for each timepoint of the episode (we set timepoints to match the acquisition times of the 1976 fMRI volumes collected as participants viewed the episode).

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/analysis_schematic_bw}
\caption{\small \textbf{Methods overview.} We used hand-annotated descriptions of each moment of video to fit a topic model.  Three example video frames, and their associated descriptions are displayed (top two rows).  Participants later recalled the video (in the third row, we show example recalls of the same three scenes from one participant).  We used the topic model (fit to the annotations) to estimate topic vectors for each moment of video and each sentence the participants recalled.  Example topic vectors are displayed in the bottom row (green: video annotations; blue: example participant's recalls).  Three topic dimensions are shown (the highest-weighted topics for each of the three example scenes, respectively).  We also show the 5 highest-weighted words for each topic.  Figure~\topics~provides a full list of the top 10 words from each of the discovered topics.}
\label{fig:schematic}
\end{figure}

The topics we found were heavily character-focused (e.g., the top-weighted word in each topic was nearly always a character) and could be roughly divided into themes related to primarily Sherlock-focused; primarily (John) Watson-focused (Watson is Sherlock's close confidant and assistant); and themes that involved Sherlock and Watson interacting (Fig.~\topics).  Several of the topics also appeared highly similar, which we hypothesized might allow us to distinguish between subtle narrative differences (if the distinctions between those overlapping topics were meaningful).  The topic vectors for each timepoint were \textit{sparse}, in that only a small number (usually one or two) topics tended to be ``active'' in any given timepoint (Fig.~\ref{fig:model}A).  Further, the dynamics of the topic activations appeared to exhibit \textit{persistance} (i.e., given that a topic was active in one timepoint, it was likely to be active in the following timepoint) along with \textit{occasional rapid changes} (i.e., occasionally topics would appear to spring into or out of existence).  These two properties of the topic dynamics may be seen in the block diagonal structure of the timepoint-by-timepoint correlation matrix (Fig.~\ref{fig:model}B).  Following \cite{BaldEtal17}, used a Hidden Markov Model (HMM) to identify the \textit{event boundaries} where the topic activations changed rapidly (i.e., at the boundaries of the blocks in the correlation matrix; event boundaries identified by the HMM are outlined in yellow).  Part of our model fitting procedure required selecting an appropriate number of ``events'' to segment the timeseries into.  We used an optimization procedure to identify the number of events that maximized within-event stability while also minimizing across-event correlations (see \textit{Methods} for additional detail).  To create a stable ``summary'' of the episode, we computed the avereage topic vector within each event (Fig.~\ref{fig:model}C).

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recalls.} All panels: darker colors indicate greater values; range: [0, 1].  \textbf{A.} Topic vectors ($K = 100$) for each of the 1976 video timepoints.  \textbf{B.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel A.  Event boundaries detected by the HMM are denoted in yellow (34 events detected).  \textbf{C.} Average topic vectors for each of the 34 video events. \textbf{D.} Topic vectors for each of 294 sentences spoken by an example participant while recalling the video.  \textbf{E.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel D. Event boundaries detected by the HMM are denoted in yellow (27 events detected).  \textbf{F.} Average topic vectors for each of the 27 recalled events from the example participant.  \textbf{G.} Correlations between the topic vectors for every pair of video events (Panel C) and recalled events (from the example participant; Panel F).  \textbf{H.} Average correlations between each pair of video events and recalled events (across all 17 participants).  To create the figure, each recalled event was assigned to the video event with the most correlated topic vector.}
\label{fig:model}
\end{figure}

@@

% characterize gist of movie and recalls
% x1. movie: fit topic models (add note on dynamic topic models).  fig 1: what do the topics reflect (top words from each topic), example with a few scenes.
% x2. notice block diagonal structure, rather than smooth transitions; motivates HMM approach.  Note about selecting K
% 3. question: do recalls exhibit same structure?  apply topic model to recalls.  note that by projecting everything into a common topic space, the thematic content of each moment of movie/recall can be compared, even though the mix of words used to describe any given event or moment may not be consistent across people (or across annotations/recall)
% 4. recalls exhibit blocky structure, but at different scales.  motivates fitting K separately by person
% 5. match up events (correlations)
% 6. now we can use classic memory analyses to examine structure of recall.  motivate and reference PFR, lag-CRP, and SPC.  but this does not tell us about structure in the *content* of what people remembered, or whether people's recalls are capturing the hierarchical temporal structure of the original movie.
% 7. topic trajectories as a new way of characterizing recall.  assess agreement using rayleigh test.  also examine shapes of each subject's trajectory to show match.  key insight: everyone captures the general shape, but at different "sampling resolutions"
% 8. we can also project back onto the original (word) features to understand more about what any particular part of the trajectory reflects (wordle fig)
% 9. the above analyses tell us about how the trajectory of thematic content in the episode is both preserved and transformed in participants' memories.  we next carried out analyses to understand the neural bases of these phenomena.  analysis 1: which structures are sensitive (during viewing) to the original episode's topic trajectory?  answer: default mode.  anlaysis 2: which structure are sensitive (during viewing) to how the participants will later recall the episode?  answer: vmPFC, default mode.  (potential role of vmPFC in reflecting participant-specific memory transformations and processing)

@@
